### 2018-11-07 2019/naacl/naaclhlt2019.tex
add bib

%
% File naacl2019.tex
%
%% Based on the style files for ACL 2018 and NAACL 2018, which were
%% Based on the style files for ACL-2015, with some improvements
%%  taken from the NAACL-2016 style
%% Based on the style files for ACL-2014, which were, in turn,
%% based on ACL-2013, ACL-2012, ACL-2011, ACL-2010, ACL-IJCNLP-2009,
%% EACL-2009, IJCNLP-2008...
%% Based on the style files for EACL 2006 by 
%%e.agirre@ehu.es or Sergi.Balari@uab.es
%% and that of ACL 08 by Joakim Nivre and Noah Smith

\documentclass[11pt,a4paper]{article}
\usepackage[hyperref]{naaclhlt2019}
\usepackage{times}
\usepackage{latexsym}

\usepackage{url}

%\aclfinalcopy % Uncomment this line for the final submission
%\def\aclpaperid{***} %  Enter the acl Paper ID here

%\setlength\titlebox{5cm}
% You can expand the titlebox if you need extra space
% to show all the authors. Please do not make the titlebox
% smaller than 5cm (the original size); we will check this
% in the camera-ready version and ask you to change it back.

\newcommand\BibTeX{B{\sc ib}\TeX}

\title{Sparse Word Embedding for Multi-domain Machine Translation}

\author{Minh Quang Pham\\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  {\tt email@domain} \\}

\date{}

\begin{document}
\maketitle
\begin{abstract}
\end{abstract}

\section{Introduction}

\section{Related Work}
Multi-domain Machine Translation has largely interested NLP community by its promising application in industry and its relation to fundamental problem in theory. Researchers have proposed a large range of techniques from Data centric methods to Model centric methods \cite{C18-1111}; \cite{P17-2061}. Data centric methods aim to augment in-domain training data by using data selection techniques such as 
The problem is recently investigated by several interesting works presented in EMNLP 2018 such as \cite{D18-1039}; \cite{D18-1041}. \cite{D18-1041} create in the encoder Domain-specific gate $g^r_i$and Domain-shared gate $g^s_i$ which are generated from domain-specific and domain-shared semantic representations of source sentence $E_r(x)$ and $E_s(x)$ respectively and which select information from units of hidden states $h_i$ by elementwise product $h^r_i = g^r_i \odot h_i$; $h^s_i = g^s_i \odot h_i$. $h^r_i$ and $h^s_i$ will be fed to Domain-specific and Domain-shared attentional mechanisms respectively. While introducing new features in the architecture, \cite{D18-1041} introduces also new objective which is sum of word-level weighted MT objective and objectives of Domain-classifier in source side and target side and Adversarial Domain-classifier in source side. The author have very good approach to well seperate Domain-shared information flow and Domain-specify information flow and determine their contributions to the inference that mitigates the catastrophic interference happens in the network during forward step.

\section{The Model}
There is not any changes in the architecture of Neural Machine Translation network except the construction of word embedding.

\section{Experiments}

\subsection{Corpora}

\subsection{NMT engine}

\subsection{Results}

\section{Conclusions}

\section*{Acknowledgments}

\bibliography{naaclhlt2019}
\bibliographystyle{acl_natbib}

\end{document}