### 2020-06-03 2020/phd_midterm_report/resume_phd.tex

\documentclass[11pt, letterpaper]{article}
\usepackage[left=2cm,right=2cm,top=1.5cm,bottom=1.5cm]{geometry}
\usepackage{enumitem}
\usepackage[utf8]{inputenc}
\usepackage{hyperref}
\begin{document}
\setlength{\belowdisplayskip}{3pt} \setlength{\belowdisplayshortskip}{3pt}
\setlength{\abovedisplayskip}{3pt} \setlength{\abovedisplayshortskip}{3pt}
\setlist{nosep}
\setlength{\parskip}{0.1cm}
\setlength{\parindent}{1em}
\section*{Minh Quang PHAM}
\begin{itemize}
	\item Formation: 
		\begin{itemize}
			\item Diplôme d'ingénieur polytechnicien, Ecole Polytechnique (2013-2016)
			\item Diplôme d'ingénieur, Telecom ParisTech (2016-2018).
			\item Master 2, Data Science, Paris-Saclay (2016-2017).
		\end{itemize}  
	\item Doctorat: Computer Science.
	\item Contract CIFRE  with Systran Group.
	\item Ecole doctorale: Sciences et Technologies de l'Information et de la Communication.
	\item Date of the first inscription: 1 November 2018.
	\item Date expected for the defense: 
	\item Supervisor: François YVON.
	\item Topic: Neural Machine Translation in Multi-Domain Context.
\end{itemize}
\section*{Publications}
\begin{itemize}
	\item Elise Michon, Minh Quang Pham, Josep M Crego, Jean Senellart. Neural Network Architectures for Arabic Dialect Identification. Proceedings of the Fifth Workshop on NLP for Similar Languages, Varieties and Dialects (VarDial 2018). \url{https://www.aclweb.org/anthology/W18-3914}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart, François Yvon. Fixing Translation Divergences in Parallel Corpora for Neural Machine Translation. In Proceedings of the 2018 Conference on Empirical Methods in Natural Language Processing, Brussels, Belgium, October 31 - November 4, 2018 , (pp. 2967$-$2973). \url{https://www.aclweb.org/anthology/D18-1328}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart. SYSTRAN Participation to the WMT2018 Shared Task on Parallel Corpus Filtering. Proceedings of the Third Conference on Machine Translation: Shared Task Papers (pp. 934$-$938). \url{https://www.aclweb.org/anthology/W18-6485}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart, François Yvon. Generic and Specialized Word Embeddings for Multi-Domain Machine Translation. In Proceedings of the 16th International Workshop on Spoken Language Translation, IWSLT, (p. 9p). Hong-Kong, CN. \url{https://hal.archives-ouvertes.fr/hal-02343215/file/IWSLT2019_paper_10.pdf}
	\item Minh Quang Pham, Josep M Crego, Jean Senellart. SYSTRAN @ WAT2019: Russian-Japanese news commentary task. In Proceedings of the 6th Workshop on Asian Translation, (pp. 189–194). Hong Kong, China: Association for Computational Linguistics. \url{https://www.aclweb.org/anthology/D19-5225}
\end{itemize}
\subsection*{Submissions}
\begin{itemize}
	\item Minh Quang PHAM, Josep CREGO, François YVON. Revisiting Multi-Domain Machine Translation. Submitted to Transactions of the Association for Computational Linguistics.
\end{itemize}
\section*{Phd courses (38 hours)}
\subsection*{Completed courses}
\begin{itemize}
	\item Mouvement Brownien et calculs stochastique (30 hours) (M2 Mathématique Aléatoire, Paris Sud/ Paris Saclay)
	\item Séquenceur Slurm pour accéder à des clusters de calcul (LRI) (8 hours)
	\item "Formation traverse" is exempted.
\end{itemize}
\subsection*{On-going courses}
\begin{itemize}
	\item Ethics \& STICs (session 2 - 2020).
\end{itemize}
\end{document}