### 2020-08-20 2020/tacl2020/resubmission-paper2327.tex

%% TACL resubmission-pkg.tex v1.1, Michael Collins and Lillian Lee
% fixed page-size handling (thanks to Ion Madrazo), added anonymous authors, simplified.
%

\documentclass[11pt,a4paper]{article}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% INPUTS FOR THE NEW PAPER -- AUTHORS, CHANGE APPROPRIATELY
\newcommand{\subtitle}{Revisiting Multi-Domain Machine Translation",}
\title{Resubmission of TACL \#2327, \\ \subtitle.   \\}

\newcommand{\coverfile}{rebuttal.pdf}  %% name of anonymized cover-letter file
\newcommand{\resubmission}{multidomain.pdf}  %% revised version
\newcommand{\origdecision}{decision.pdf}  %% name of anonymized original-decision-letter file

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{hyperref}
\hypersetup{
    colorlinks=true, %set true if you want colored links
    linktoc=all,     %set to all if you want both sections and subsections linked
    linkcolor=blue,  %choose some color if you want links to stand out
}

\usepackage{times}
\usepackage{pdfpages}
\pagestyle{plain}

\author{(Anonymous TACL resubmission)}
\setlength{\footskip}{55pt}

\begin{document}
\maketitle

\tableofcontents

\section{Author(s) cover letter responding to the original reviews} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\coverfile}
\section{Revised submission} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\resubmission}
\section{Original decision letter and reviews} Starts on next page.
\includepdf[pages=-,fitpaper=true]{\origdecision}

\end{document}