### 2021-04-05 2021/thesis/parts/state_of_the_art.tex

\part{State of the art}
\chapter{Neural Machine Translation}
\section{Recurrent Neural Network models}
\subsection{GRU, LSTM}
\subsection{Attentional mechanism}
\section{Transformer models}
\subsection{self-attention, cross-attention layers}
\subsection{Variants}
\chapter{Data centric domain adaptation}
\chapter{Feature centric domain adaptation}
\chapter{Model centric domain adaptation}