### 2021-05-12 2021/emnlp/auto_cl_mdmt.tex
fix some lambda, resolve fixmes

% This must be in the first 5 lines to tell arXiv to use pdfLaTeX, which is strongly recommended.
\pdfoutput=1
% In particular, the hyperref package requires pdfLaTeX in order to break URLs across lines.

\documentclass[11pt]{article}

% Remove the "review" option to generate the final version.
\usepackage[review]{emnlp2021}

% Standard package includes
\usepackage{times}
\usepackage{latexsym}

% For proper rendering and hyphenation of words containing Latin characters (including in bib files)
\usepackage[T1]{fontenc}
% For Vietnamese characters
% \usepackage[T5]{fontenc}
% See https://www.latex-project.org/help/documentation/encguide.pdf for other character sets

% This assumes your files are encoded as UTF8
\usepackage[utf8]{inputenc}
\renewcommand{\UrlFont}{\ttfamily\small}
\usepackage{graphicx}
\usepackage{grffile}
\usepackage{multirow}
\usepackage{xcolor,colortbl}
\usepackage{amsmath}
%\usepackage{fdsymbol}
\usepackage{amssymb}
\usepackage{makecell}
\usepackage[super]{nth}
\usepackage{arydshln}

\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{regexpatch}
\usepackage{subcaption}
% This is not strictly necessary, and may be commented out,
% but it will improve the layout of the manuscript,
% and will typically save some space.
\usepackage{microtype}

\usepackage{url}
\usepackage{longtable}
\usepackage{tikz}
\usetikzlibrary{calc}
\usepackage[draft]{todo}
\usepackage[normalem]{ulem}
\usepackage{xspace}
\usepackage{float}

\usepackage{pgfplots}
\usepackage{pgfplotstable}

\let\oldbibitem\bibitem
\def\bibitem{\vfill\oldbibitem}

\usepackage{soulutf8}
\usepackage{tabularx}

\usepackage{multirow}
\usepackage{adjustbox}
\usepackage{caption}
\usepackage{subcaption}

\newcommand{\fyTodo}[1]{\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyTodostar}[1]{\Todo*[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyDone}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}
\newcommand{\fyFuture}[1]{\done[FY]\Todo[FY:]{\textcolor{red}{#1}}}
\newcommand{\fyDonestar}[1]{\done[FY]\Todo[FY:]{\textcolor{orange}{#1}}}

\newcommand{\revision}[1]{\textcolor{red}{#1}}
\newcommand{\revisiondel}[1]{}
\newcommand{\src}{\ensuremath{\mathbf{f}}} % source sentence
\newcommand{\trg}{\ensuremath{\mathbf{e}}} % target sentence
\newcommand{\domain}[1]{\texttt{\textsc{#1}}}
\newcommand{\system}[1]{\texttt{{#1}}}

\newcommand{\vlambda}{\ensuremath{\boldsymbol\lambda}\xspace} % parameters vector for a distribution
\newcommand{\vtheta}{\ensuremath{\boldsymbol\theta}\xspace} % parameters vector for a distribution
\newcommand{\vpsi}{\ensuremath{\boldsymbol\psi}\xspace} % parameters vector for a distribution
\newcommand{\indic}[1]{\ensuremath{\mathbb{I}(#1)}}
% \newcommand{\SB}[1]{\textcolor{green}{#1}}
% \newcommand{\SW}[1]{\textcolor{red}{#1}}
\newcommand{\SB}[1]{\textbf{#1}}
\newcommand{\SW}[1]{\underline{#1}}
% limits underneath
\DeclareMathOperator*{\argmin}{arg\,min}
\DeclareMathOperator*{\argmax}{arg\,max}
\renewcommand\textfraction{.1}
\renewcommand\floatpagefraction{.95}
\newcommand{\sbcl}[2]{{\scriptsize #1 \hfill $|$ \hfill  #2}}
\usepackage{multirow}
\usepackage{adjustbox}

\title{Multi-Domain Adaptation in Neural Machine Translation with Dynamic Sampling Strategies}

\author{First Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\\And
  Second Author \\
  Affiliation / Address line 1 \\
  Affiliation / Address line 2 \\
  Affiliation / Address line 3 \\
  \texttt{email@domain} \\}

\date{}

\begin{document}
\maketitle
\setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt}
\begin{abstract}
  Building effective Neural Machine Translation models often imply accommodating diverse sets of heterogeneous data so as to optimize performance for the domain(s) of interest. Such multi-source / multi-domain adaptation problems are typically approached through instance selection or reweighting strategies, based on a static assessment of the relevance of training instances with respect to the task at hand. In this paper, we study dynamic data selection strategies that are able to automatically re-evaluate the usefulness of data samples and to evolve a data selection policy in the course of training. Based on the results of multiple experiments, we show that such methods constitute a generic framework to automatically and effectively cater to a variety of real-world situations, from supervised multi-source domain adaptation to multi-domain learning and unsupervised domain adaptation. 
\end{abstract}

\section{Introduction}\label{sec:intro}
A typical setting in machine translation (MT) is to collect the largest possible collection of parallel data for the chosen language pair, with the intent to achieve optimal performance for a task of interest. In such situations, the training data distribution is opportunistic, while the test data distribution is chosen and fixed; a key component in training is then to mitigate the detrimental effects of a possible mismatch between these distributions. Single-source and multi-source\footnote{In this context, multi-source refers to simultaneously having multiple domains to adapt from and should not be confused with multi-source translation, where several \emph{source languages} are considered.} domain adaptation (DA) is a well-studied instance of this setting (see \citet{Chu2017comparison} for a review), and so is multi-domain (MD) learning \citep{Chu18multilingual,Zeng18multidomain,Jiang19multidomain,Pham21revisiting}. A related situation is multilingual machine translation \citep{Firat16multiway,Ha16towards,Johnson17google,Arivazhagan19massively}\fyTodo{Fix ref}\revision{cite also \citep{Shen21source}}, where the heterogeneity of training data not only corresponds to variations in topic, genre or register, but also in language.

This problem is often approached by \emph{static} instance selection or re-weighting strategies, where the available training data is used in proportion to its relevance for the chosen test conditions \citep{Moore10selection,Axelrod11domain}. Finding the optimal balance of training data is however a challenging task due, for instance, to the similarity between domains/languages, but also due to regularization effects of out-of-domain data. It may also be suboptimal, as some target domains or languages might be easier to train than others. For example, improving the performance of the MT system in one domain will often hurt that of another \citep{Vanderwees17dynamic, Britz17mixing}, and improving model generalization across all domains \citep{koehn18findings} may not achieve optimily for any particular domain.

Several recent proposals
% \citep{Vanderwees17dynamic,Graves17automated,platanios19competence,Zhang19curriculum,Kumar19reinforcement,Wang20learning-multi,Wang20balancing,Wang20learning-multi}
have explored ways to instead consider \emph{dynamic} data selection and sampling strategies that surpass static strategies. Notably,  \citet{Vanderwees17dynamic,Zhang19curriculum} construct a static curriculum, while\citet{Graves17automated,platanios19competence,Kumar19reinforcement,Wang20balancing,Wang20learning-multi} build curricula that automatically adapt to the training data.

In this paper, we contribute to this line of research in several ways. First, we propose a novel framework (\emph{Multi-Domain Automated Curriculum}, MDAC for short) that simultaneously accounts for the domain adaptation and the multi-domain adaptation problems. Second, we propose a novel method of building an automated curriculum for training Multi-Domain NMT models. We evaluate our method in various single-domain and multi-domain adaptation settings. We compare our method with several contrasts including the work of \citet{Zhang19curriculum} and \citet{Wang20balancing}.
% From now on, we name our method MDAC which is brief for .
Based on these experimental results, our main conclusions are that (a) using MDAC often yields overall performance that is as good as the standard finetuning strategy for domain adaptation; (b) MDAC can effectively handle a variety of test situations, from targeting one single domain to the full multi-domain scenario. % distributions

\section{Learning with multiple data sources} \label{sec:mdmt}

We conventionally define a domain $d$ as a distribution $\mathcal{D}_d(x)$ over some feature space $\mathcal{X}$ that is shared across domains \citep{Pan10asurvey}: in machine translation, $\mathcal{X}$ is the representation space for input sentences; each domain corresponds to a specific source of data, and may differ from the other data sources in terms of textual genre, thematic content \citep{Chen16guided,Zhang16topicinformed}, register \cite{Sennrich16politeness}, style \cite{Niu18multitask}, etc. Translation in domain $d$ is formalized by a translation function $h_d(y|x)$ pairing sentences in a source language with sentences in a target language $y \in \mathcal{Y}$. $h_d$ is usually assumed to be deterministic (hence $y = h_d(x)$), but might differ from one domain to the other.

It is usual in MT to opportunistically collect training samples from several domains, which means that training instances are distributed according to a mixture $\mathcal{D}^s$ such that $\mathcal{D}^s(x) = \sum_{d=1}^{n_d} \lambda^{s}(d) \mathcal{D}_d(x)$, with $\{\lambda^{s}(d), d=1 \dots n_d\}$ the mixture weights satisfying $\sum_d \lambda^{s}(d)=1$.

The main challenge is then to make the best of this heterogeneous data, with the aim to achieve the optimal performance for the intended test conditions. These might correspond to data from just one of the training domains, as in standard supervised (multi-source) domain adaptation; a more difficult case is when the test data is from one domain unseen in training (unsupervised domain adaptation); in multi-domain adaptation finally, the test distribution is itself a mixture of domains, some of which may also be observed in training.  Without loss of generality, one may then assume that the test distribution takes the form $\mathcal{D}^{t}(x) = \sum_d \lambda^{t}(d) \mathcal{D}_d(x)$ - with only one non-null  component in the case of pure domain adaptation.
These various settings are illustrated Figure~\ref{fig:mdmt-lambdas}.
\begin{figure}[h]
  \centering
  \includegraphics[width=0.48\textwidth]{mdmt-lambdas}
  \caption{Training and testing with distribution mismatch. We consider just three domains, and represent vectors of mixture weights $\vlambda^{s}$ and $\vlambda^{t}$ in the 3-dimensional simplex. Training with weights in (a) and testing with weights in (c) is supervised multi-source domain adaptation to domain~2 ($d_2$), while (b)-(c) is the unsupervised version, with no training data from $d_2$; training with weights in (a) and testing with weights in (d) is multi-domain learning, also illustrated with configurations (a)-(e) (training domain $d_1$ is not seen in test), and (b)-(d)  (test domain $d_2$ is unseen in training).}\label{fig:mdmt-lambdas}
\end{figure}

These situations have been amply documented from a theoretical perspective (eg.\ \cite{Mansour09multiple,Mansour09domainadaptation,Hoffman18algorithms}). A general recommendation in the DA setting is to adjust the sampling distribution used to optimize the system so as to compensate for the mismatch between $\mathcal{D}^s(x)$ and $\mathcal{D}^t(x)$. This can be approximated by reweighting instances, or more conveniently domains, which are selected during training with a probability $\lambda^{l}(d)$, with $\lambda^{l}(d) \neq \lambda^{s}(d)$.

A more practical standard approach to supervised DA is \emph{finetuning} \cite{Luong15stanford,Freitag16fast}, where $\vlambda^{l}$ is allowed to vary during learning. With our notations, this approach amounts to first learning an initial parameter value with all the data ($\forall d, \lambda^{l}(d) = \lambda^{s}(d)$), then to continue training with only batches from the test domain $d_t$ ($\lambda^{l}(d) = \indic{d = d_t}$) with $\indic{A}$ the indicator function for predicate $A$. Note that this strategy is potentially suboptimal, as some out-of-domain samples may contribute to the final performance due to eg.\ domain overlap. Optimizing the learning distribution in multi-domain settings is even more challenging, as the learner has to best take advantage of potential domains overlaps, and also of the fact that some domains might be easier to learn than others.\fyTodo{How to measure this?} \revision{Answer: The graphic of \system{Mixed-0} shows that with the same amount of data(batches) the convergences of the domains vary significantly}

% === The part on multidomain training has been moved to the end.

\section{Multi-Domain Automated Curriculum } \label{sec:mdac}
\subsection{Basic principles}
Assuming training data in each of the $n_d$ domains $d_1 \dots d_{n_d}$, we denote the size of the training corpus from domain $d$ as $N^{s}_d$, and $N^{s} = \sum_d N^{s}_d$ is the total number of training samples. We use $\widehat{\mathcal{D}^l_d}$ and $\widehat{\mathcal{D}^t _d}$ to denote the empirical train and test distributions for domain $d$, and $\widehat{\mathcal{D}^{u}}(x;\lambda^{u}) = \sum_{d} \lambda^{u}(d) \widehat{\mathcal{D}^{u}_d}(x)$ for $u\in\{l,t\}$. In our setting,  $\vlambda^t$, and hence $\widehat{\mathcal{D}^t}(x;\vlambda^t)$ are fixed and predefined, approximated with an equivalent number of development corpora.

% Our method
MDAC constructs an adaptative training distribution $\lambda^{l}$ that optimizes the data selection policy along with the training of the NMT model. We parameterize $\lambda^{l}$ by a differentiable function $\lambda^l(\psi)$. We divide the training into many short sessions; in each session $t$, the model is trained with a static data distribution $\lambda^{l}(\psi_t)$. After one learning session, we update the data distribution using the REINFORCE algorithm of \citet{Williams92simple}. The evolution of $\psi$ is thus defined by:
\begin{align*}
\psi_{t+1} &= \psi_t + \mathbf{lr}_{1} * \displaystyle{\mathop{\sum}_{d=1}^{n_d}} R(d) * \frac{\partial \lambda^l(d;\psi_t)}{\partial \psi}, \\
\end{align*}
\begingroup
\allowdisplaybreaks
where reward $R(d)$ is computed as:
\begin{align*}
  R(d) = J^t(\theta_{t+k},\vlambda^t) - J^t(\theta_t,\vlambda^t),
\end{align*}
and where we also define:
\begin{equation}
\begin{array}{rcl}
\theta_{t+i} &=& Update\big(\theta_{t+i-1},[x^i_j,y^i_j]_{j=1}^N\big) \\ \nonumber
x^i_j, y^i_j &\sim& \widehat{D^l_d}(x) \\
J^t(\theta,\lambda_t) &=& \displaystyle{\mathop{\sum}_{d=1}^{n_d}}\lambda^t(d)\displaystyle{\mathop{\sum}_{x^t_d,y^t_d \in \widehat{D^t_d}}} l(\theta,x^t_d,y^t_d).
\end{array}
\end{equation}
\endgroup
% \theta_{t+i-1} - \mathbf{lr_2} \frac{1}{N}\displaystyle{\mathop{\sum}_{j=1}^N}\frac{\partial l(\theta_{t+i-1}, x^i_j,y^i_j)}{\partial \theta}
In these equations, $N$ denotes the size of a batch; $\mathbf{lr}_{1}$ is the learning rate of the sampling distribution; $\mathbf{lr_2}$ is the learning rate of the NMT system; $l(\theta,x,y)$ is the loss of the NMT model on sample $(x,y)$; $J^t(\theta,\lambda_t)$ is the weighted loss aggregated over $n_d$ dev-sets corresponding to the $n_d$ domains.

To compute the reward $R(d)$ of using data from domain $d$ to train the model, we simulate $k$ training steps from current checkpoint by training the model with $k$ batches sampled from $D^l(d)$ and compute the gain of the weighted dev-loss. This computation is inspired by the target prediction gain in \citep{Graves17automated}. However, \citep{Graves17automated} used accumulated gains from the past as reward, while we predict the usefulness of each domain for improving the future performance of the system given its current state. This is achieved by simulating a round of training with only the data from one domain. Furthermore, the choice of the parameterization of the sampling distribution in our method differs from that of \citet{Graves17automated}, although this choice is empirical.

The work of \citet{Wang20balancing} is also related: it is based on the Bi-level Optimization framework, which aims to find an optimal static distribution $\lambda^{l}$ that will result in the best NMT model with respect to a given target dev set at the end of training. These authors also derive a similar form of update for $\psi$. However, their reward is the cosine similarity between the gradient computed with the training data from one domain and the gradient computed with the dev set. We compare this approach with ours in the experiment section.

\subsection{MDAC for (multi) domain adaptation}
The setting developed in previous sections is quite general and can in principle accommodate the variety of situations mentioned above, and many more: basic domain adaption, multi-domain adaptation with various target distributions, possibly including domains unseen in training. In our experiments, we would like to better assess the true potential of MDAC is these settings and seek to experimentally answer the following questions:
\begin{itemize}
\item is MDAC a viable alternative to conventional finetuning? In particular, does it enable to better take advantage of relevant data from other domains?
\item is MDAC also a viable option in multi-domain adaptation scenarios?
\item does MDAC also enable to perform \emph{unsupervised} (multi-)domain adaptation? \fyDone{TBContinued}
\end{itemize}
These questions are further explored in Section~\ref{sec:results}. We now turn to our experimental conditions.

\section{Experimental settings} \label{sec:exp}
\subsection{Data and metrics \label{ssec:corpora}}
We experiment with translation from English into French and use texts initially originating from 6~domains, corresponding to the following data sources: the UFAL Medical corpus V1.0 (\domain{med})\footnote{\url{https://ufal.mff.cuni.cz/ufal_medical_corpus}. We only use the in-domain (medical) subcorpora: PATR, EMEA, CESTA, ECDC.}; the European Central Bank corpus (\domain{bank}) \cite{Tiedemann12parallel}; the JRC-Acquis Communautaire corpus (\domain{law}) \cite{Steinberger06acquis}; documentations for KDE, Ubuntu, GNOME and PHP from the Opus collection \cite{Tiedemann09news}, merged in a \domain{it}-domain; TedTalks (\domain{talk}) \cite{Cettolo12wit}, and the Koran (\domain{rel}). Complementary experiments also use v12 of the News Commentary corpus (\domain{news}). Most corpora are available from the Opus website.\footnote{\url{http://opus.nlpl.eu}} These corpora were deduplicated and tokenized with in-house tools; statistics are in Table~\ref{tab:Corpora}. To reduce the number of types and build open-vocabulary systems, we use Byte-Pair Encoding \cite{Sennrich16BPE} with 30,000 merge operations on a corpus containing all sentences in both languages.\fyDone{Add \# number of tokens, also specificity ?}%

We randomly select in each corpus a development and a test set of 1,000 lines and keep the rest for training. Validation sets are used to chose the best model according to the average BLEU score \citep{Papineni02bleu}.\footnote{We use truecasing and the \texttt{multibleu} script.}\fyDone{A word about meta-parameter settings} Significance testing is performed using bootstrap resampling \cite{Koehn04statistical}, implemented in compare-mt\footnote{\url{https://github.com/neulab/compare-mt}} \citep{Neubig19compare-mt}. We report significant differences at the level of $p=0.05$.\fyDone{Fix correct p value.}\fyTodo{Pourquoi pas sacre BLEU?}

\begin{table*}[htbp]
  \centering
  \begin{tabular}{|l|ccccccc|} %*{4}{|r|}}
    \cline{2-8} 
    %\multicolumn{4}{|l|}{Vocab size - En: 30,165, Fr: 30,398}\\
    \multicolumn{1}{c|}{} & \multicolumn{1}{c}{\domain{med}} & \multicolumn{1}{c}{\domain{law}} & \multicolumn{1}{c}{\domain{bank}} & \multicolumn{1}{c}{\domain{it}} & \multicolumn{1}{c}{\domain{talk}} & \multicolumn{1}{c}{\domain{rel}} & \multicolumn{1}{c|}{\domain{news}} \\
    \hline 
    \# lines & 2609 (0.68) & 501 (0.13) & 190 (0.05) & 270 (0.07) & 160 (0.04) & 130 (0.03) & 260 (0) \\
    \# tokens &  133 / 154  &  17.1 / 19.6 &  6.3 / 7.3 &  3.6 / 4.6 &  3.6 / 4.0 &  3.2 / 3.4 & 7.8 / 9.2   \\
    \# types & 771 / 720 & 52.7 / 63.1 & 92.3 / 94.7 & 75.8 / 91.4 & 61.5 / 73.3 & 22.4 / 10.5 & - \\
    \# uniq & 700 / 640 & 20.2 / 23.7 & 42.9 / 40.1 & 44.7 / 55.7 & 20.7 / 25.6 & 7.1 / 2.1 & - \\
    \hline
  \end{tabular}
  \caption{Corpora statistics: number of parallel lines ($\times 10^3$) and proportion in the training domain mixture (which does not contain \domain{news}), number English and French tokens ($\times 10^6$), number English and French types ($\times 10^3$), number of types that only appear in a given domain ($\times 10^3$). \domain{med} is the largest domain, containing almost 70\% of the sentences, while \domain{rel} is the smallest, with only 3\% of the data.
  }
\label{tab:Corpora}
\end{table*}

\subsection{Baseline systems \label{ssec:baseline}}
Our baselines are standard for multi-domain systems.\footnote{We however omit domain-specific systems trained only with the corresponding subset of the data, which are always inferior to the mix-domain strategy \cite{Britz17mixing}.} Using Transformers \cite{Vaswani17attention} implemented in OpenNMT-tf\footnote{\url{https://github.com/OpenNMT/OpenNMT-tf}} \cite{Klein17opennmt}, we build the following systems:

\begin{itemize}
\itemsep0em 
\item Generic models trained with various predefined mixtures of the training data taking the form:
\begin{align} \label{mixture:trn}
\lambda_{\alpha}(d) = \frac{q_d^{\alpha}}{\displaystyle{\mathop{\sum}_{d=1}^{n_d}q_d^{\alpha}}} &&
q_d = \frac{\mid N^{s}_d \mid}{\displaystyle{N^{s}}} % \mathop{\sum}_{i=1}^K\mid D_i \mid}}
\end{align} \setlength{\abovedisplayskip}{2pt}
\setlength{\belowdisplayskip}{2pt} 
with $\alpha \in [0,0.25,0.5,0.75,1.0]$. These systems are denoted \system{Mixed-$\alpha$} below. \system{Mixed-$0$} uses a uniform domain distribution, \system{Mixed-$1.0$} uses the observed domain distribution.
\item finetuned models \citep{Luong15stanford,Freitag16fast}, based on the \system{Mixed-$1.0$} system, further trained on each domain for at most 50~000 iterations, with early stopping when the dev BLEU stops increasing for 5 successive iterations.\fyDone{for ?? iterations} The full finetuning (\system{FT-Full}) procedure may update all the parameters of the initial generic model, resulting in six systems, each adapted for one domain, with no parameter sharing across domains.

\item Our comparison of multi-domain systems\fyTodo{CL is only dynamic ?} includes baselines with fixed data mixtures corresponding to $\vlambda^l \in \big[ \vlambda_0, \vlambda_{0.25}, \vlambda_{0.5}, \vlambda_{0.75}, \vlambda_{1.0}\big]$, and our reimplementations of recent proposals from the literature: Curriculum Learning (CL) of \citet{Zhang19curriculum} and Differential Data Selection (DDS) of \citet{Wang20balancing}.
\end{itemize}

All models use embeddings and hidden layers sizes of dimension~512. Transformer models contain 8~attention heads in each of the 6+6 layers; the inner feedforward layer contains 2048 cells. The adapter-based systems (see below) additionally use an adaptation block in each layer, composed of a 2-layer perceptron, with an inner $\operatorname{ReLU}$ activation function operating on normalized entries of dimension~1024. 
Training uses batches of~12,288 tokens, Adam with parameters $\beta_1=0.9$, $\beta_2= 0.98$, Noam decay ($warmup\_steps=4000$), and a dropout rate of $0.1$ in all layers. We train all systems with 200~000 iterations.

\subsection{MDAC systems} \label{ssec:dds-sys}
The behavior of MDAC only depends on (a) the initial domain distribution at the start of training $\vlambda^{l}_{t=0}$, and (b) the targeted (dev/test) distribution $\vlambda^{t}$. We will thus report these systems as \system{MDAC($\vlambda^{l}_{t=0}$, $\vlambda^{t}$)}. We also compare our method with DDS under the same settings.

In our work, we simply parameterize the distribution $\vlambda^l$ as follows (with $\beta=2$ in all our experiments):\fyTodo{Explain why}
\begin{equation}
\lambda
^l(d;\vpsi) = \frac{\vpsi[d]^\beta}{\sum_i \vpsi[i]^\beta}. \nonumber
\end{equation}

Additional choices for the hyper-parameters of our method include the number of simulation steps $k=10$ and the learning rate $\mathbf{lr}_{data}=0.001$. We update the sampling distribution via 100 gradient descent iterations for almost all experimental settings except that for adaptation with automatic clusters (Section~\ref{ssec:clda}), where we use only 20 gradient descent iterations.\footnote{This is intended to regularize the training of $\psi$ and avoid converging towards degenerate solutions where some clusters get no training samples.}\fyDone{Why? computation?} We split the training into 100 short sessions that last 2000 training steps each. The choice of those hyper-parameters is mostly heuristic except for the learning rate $\mathbf{lr}_{data}$ which is optimized via grid search over a range of values $[0.001,0.0025,0.005]$.

\subsection{CL and DDS's re-implementation}\fyTodo{Right place for this ?}
We re-implement DDS in Tensorflow without any change in the choices of parameterization and hyper-parameters compared to original code of \citet{Wang20balancing}.\footnote{\url{https://github.com/cindyxinyiwang/multiDDS}}

We also re-implement the work of \citet{Zhang19curriculum} according to the authors' description of the algorithm. For each domain adaptation experiment, we combine the training data of all other domains into one corpus, then compute the cross-entropy difference score of each source sentence of this combined dataset. We then sort and split the corpus into 9 shards and execute curriculum learning with 10 shards, using the in-domain corpus as the first shard.

\subsection{Experimental tasks}
We evaluate and compare our method with multiple baselines in 5~tasks including: domain adaptation; multi-domain adaptation, bi-domain adaptation, unseen domain adaptation, and unsupervised domain adaptation.

In the domain adaptation task, given the data from 6 domains (\domain{med}, \domain{bank}, \domain{law}, \domain{it}, \domain{talk}, \domain{rel}), we aim to build distinct expert NMT models for each domain. In the bi-domain adaptation task, given again same 6 domains, we only focus on adapting only to 2 (known) domains.

In the multi-domain adaptation task, given the same 6 domains, we aim to build one single NMT model that would perform optimally assuming a uniform distribution of domains during the test.

In the fourth experiment (few shot adaptation)\fyTodo{or weakly supervised ?}, given training data in 6 domains and a small development set of a new domain, (\domain{News} in our case), our goal is to build an NMT model which will perform well on test data from the unseen domain.

Finally, in the unsupervised domain adaptation task, we cluster all available training data into 30 clusters using the KNN algorithm in the same way as in the work of \citet{Tars18multidomain}, then adapt these clusters to one of 6 domains given the corresponding in-domain dev set. We compare MDAC to DDS over 6 domains.

\section{Results and discussion \label{sec:results}}

\subsection{Domain Adaptation}\label{ssec:da}
In this setting, we aim to build an expert NMT model in only one domain. This setting corresponds to Dirac distribution $\vlambda^t$, in which the probability of domains is 0 except for one domain.

We initialize MDAC and DDS by 3 different distributions $\vlambda_0$, $\vlambda_1$ and $\vlambda_d$. According to Table~\ref{tab:da}, MDAC achieves the best performance with $\vlambda_0$ (uniform domain distribution) as the initial value, though $\vlambda_0$  initialization only significantly improves over $\vlambda_d$ initialization in small domains such as \domain{talk},\domain{bank} and\domain{it}. Initializing with $\vlambda_d$ seems more appropriate when targeting large domains such as \domain{med} and \domain{law}. The same conclusions also hold for DDS.

We now compare the best MDAC system (initialized with $\vlambda_{t=0} = \vlambda_0$) to full finetuning. According to Table~\ref{tab:da}, MDAC is outperformed by fine-tuning method in large domains including \domain{med} and \domain{law}. However, MDAC improves fine-tuning by approximately 0.5 BLEU in \domain{bank} and 1.0 BLEU in \domain{rel}. This indicates that for small domains, out-of-domain data can improve the generalization of the NMT model and MDAC is capable to exploit this advantage by balancing the in-domain and the out-of-domain training data instead of edging out the out-of-domain training data as in fine-tuning.

DDS also demonstrates promising results in this task, but is always outperformed by MDAC overall. The same conclusions also hold for the CL method.\fyTodo{Continue}

\begin{table*}[htbp]
  \centering \small
  \begin{tabular}{|l|*8{r|}} \hline
    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline
    FT-Full($d$) &37.7&59.2&54.5&34&46.8&90.8&53.8\\ \hline
    \hline
    \system{CL($d$)} &37.2&55.4&53.0&31.3&45.3&90.8&52.2\\ \hline
    \system{DDS($\vlambda_0, \vlambda_d$)} &36.7&56.2&54.3&33.5&46.8&91.8&53.2\\
    \system{MDAC($\vlambda_0, \vlambda_d$)} &36.7&58.2$^{**}$&55$^{*}$&33.3&46.6&91.9$^{***}$&53.6\\
     \hline
    \system{DDS($\vlambda_1, \vlambda_d$)} &36.4&56.3& 45.4& 32.6&44.5&43.1&43.1\\
    \system{MDAC($\vlambda_1, \vlambda_d$)} &37.2&55.1&51.8&32.8&44.9&85.4&51.2 \\ \hline
    \system{DDS($\vlambda_d, \vlambda_d$)} &37.0&59.6&53.7&30.2&45.7&91.6&53.0 \\
    \system{MDAC($\vlambda_d, \vlambda_d$)}&37.9&59.6&53.9&30.4&45.8&92.1&53.3 \\
    \hline
  \end{tabular}
  \caption{We report BLEU scores of each method with 6 target domains, and compare with full fine-tuning: each column corresponds to a distinct system. ($^{*}$) MDAC is significantly better than CL, fine-tuning and DDS with $p<0.05$. ($^{**}$) MDAC is significantly better than CL and DDS with $p<0.05$. ($^{***}$) MDAC is significantly better than CL, fine-tuning with $p<0.05$}
  \label{tab:da}
\end{table*}

\subsection{Multi-domain adaptation}\label{ssec:mda}

Multi-domain adaptation task requires best averaged performance of the NMT model in 6 domains. This setting corresponds to a uniform test distribution $\vlambda^t = \vlambda_0$. In this situation, the CL method of \citet{Zhang19curriculum} does not apply. We therefore only compare the performance of MDAC, DDS and several fixed training data distribution $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ (see equation~\eqref{mixture:trn}).

We initialize MDAC and DDS by 2 different distribution $\lambda_0$ and $\lambda_1$. According to Table \ref{tab:multi-da}, MDAC achieves the best performance with $\lambda_0$ initial where all domains have the same probability to be chosen. The same conclusion is also true for DDS.

MDAC outperforms static training distributions including $\big[ \lambda_0, \lambda_{0.75}, \lambda_{1.0}\big]$ with significant margin while performs slightly better than $\big[ \lambda_{0.25}, \lambda_{0.5} \big]$. This observation indicates that MDAC allows us to skip the heuristic choice for the training mixture of different domains or tasks.

The second observation is that DDS is again outperformed by MDAC with a significant margin of 0.5 BLEU on average. However, MDAC is surpassed by far with 1.8 BLEU in the domain \domain{med} by DDS.

We show in the below figure \ref{fig:sampling}, the evolution of the training mixture through training sessions. According to the figure \ref{fig:DDS} of DDS, the proportion of 6 domains converges rapidly to bi-domain mode in which only \domain{med} and \domain{rel} have significant probability. In contrast, the proportion of 6 domains computed by MDAC evolves smoothly; small domains including \domain{bank}, \domain{it}, \domain{talk} and \domain{rel} take larger proportion in the beginning then decrease their proportions while larger domains as \domain{med} and \domain{law} increase their proportions in the end where the NMT model might have been already close to the optimal performance in the smaller domains.

\begin{table*}[htbp]
  \centering \small
  \begin{tabular}{|l|*8{r|}} \hline
    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} \\ \hline \hline
    % Multidomain
    \system{Mixed-0}      & 35.3 & 54.1 & 52.5 & 31.9 & 44.9 & 89.5& 51.4 \\
    \system{Mixed-0.25} & 35.9 & 54.9 & 52.6 & 32.6 & 45 & 90.3& 51.9 \\
    \system{Mixed-0.5}   & 36.1 & 55.4 & 51.8 & 33.5 & 46.2 & 90 & 52.2 \\
    \system{Mixed-0.75} & 36.5 & 55 & 51.2 & 34 & 44.3 & 87.2& 51.7 \\
    \system{Mixed-1}     & 37.3  & 54.6 & 50.1 & 33.5 & 43.2 & 77.5& 49.4 \\
    \hline \hline
    \system{DDS($\vlambda_0, \vlambda_0$)} &37.3&55.1&51&33.5&43.4&90.8&51.8\\ 
    \system{MDAC($\vlambda_0, \vlambda_0$)}&35.5&55.5&53.7$^*$&32.3&45.6&91.1$^*$&52.3$^{**}$\\ 
    \hline \hline
    \system{DDS($\vlambda_1, \vlambda_0$)} &37.9&50.2&47.2&31.2&40.8&58.6&44.3\\
    \system{MDAC($\vlambda_1, \vlambda_0$)}&37.4&54.4&50.3&32.0&42.7&80.2&49.5\\
    \hline
  \end{tabular}
  \caption{We report the performance of our method MDAC and the work of \citet{Wang20balancing} DDS in multi-domain adaptation task. For a given line, all the columns \emph{correspond to the same multi-domain system}. ($^*$) MDAC is significantly better than \system{Mixed-$\alpha$} with $p<0.05$ and DDS. ($^{**}$) MDAC is significantly better than DDS with $p<0.05$.}
  \label{tab:multi-da}
\end{table*}

\begin{figure*}[htbp]
\begin{subfigure}{.5\textwidth}
  \centering
  % include first image
  \includegraphics[width=.8\linewidth]{DDS.png}  
  \caption{DDS}
  \label{fig:DDS}
\end{subfigure}
\begin{subfigure}{.5\textwidth}
  \centering
  % include second image
  \includegraphics[width=.8\linewidth]{MDAC.png}  
  \caption{MDAC}
  \label{fig:MDAC}
\end{subfigure}
\caption{Sampling distribution}
\label{fig:sampling}
\end{figure*}

\subsection{Bi-domain adaptation}\label{ssec:bida}

\begin{table*}[htbp]
  \centering \small
  \begin{tabular}{|l|*7{r|}} \hline
    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} \\ \hline \hline%  \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} 
    \system{MDAC($\vlambda_0, \vlambda_2$)}&36.2&-&-&-&46.1$^*$&-&\\
    \system{DDS($\vlambda_0, \vlambda_2$)}&36.5&-&-&-&44.4&- \\
    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&57.2$^*$&53.8$^*$&-&-&-& \\
    \system{DDS($\vlambda_0, \vlambda_2$)}&-&56.1&52.6&-&-&-&- \\
    \system{MDAC($\vlambda_0, \vlambda_2$)}&-&-&-&32&-&90 \\
    \system{DDS($\vlambda_0, \vlambda_2$)}&-&-&-&32.9&-&91.1 \\ \hline
  \end{tabular}
  \caption{Non empty columns correspond to the target domains. For a given line, all the non empty columns \emph{correspond to the same bi-domain system}. ($^*$) MDAC is significantly better than DDS with $p<0.05$.}
  \label{tab:bi-da}
\end{table*}
According to Table \ref{tab:bi-da}, MDAC outperforms DDS in 2 over 3 settings including adapting to \domain{med-it} and \domain{law-bank} while being surpassed in \domain{talk-rel}, showing that MDAC is capable to adapt to a variable set of target domains.

\begin{table*}[htbp]
  \centering \small
  \begin{tabular}{|l|*9{r|}} \hline
    domain \hfill $d=$ & \multicolumn{1}{c|}{\domain{ med}} & \multicolumn{1}{c|}{\domain{ law}} & \multicolumn{1}{c|}{\domain{bank}} & \multicolumn{1}{c|}{\domain{talk}} & \multicolumn{1}{c|}{\domain{ it }} & \multicolumn{1}{c|}{\domain{ rel}} & \multicolumn{1}{c|}{mean} & \multicolumn{1}{|c|}{\domain{news}} \\ \hline \hline
    \multicolumn{9}{|c|}{\sl Unsupervised domain adaptation} \\ \hline
    \system{Mixed-0}      & - & - & - & - & - & - & - & 22.4 \\
    \system{Mixed-0.25} & - & - & - & - & - & - & - & 22.7\\
    \system{Mixed-0.5}   & - & - & - & - & - & - & - & 23.6\\
    \system{Mixed-0.75} & - & - & - & - & - & - & - & 23.7\\
    \system{Mixed-1} & - & - & - & - & - & - & - & 23.5\\
    \hline \hline
    \system{MDAC($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.1 \\
     \system{DDS($\vlambda_0, \vlambda_{news}$)} &-&-&-&-&-&-&-&23.3 \\ \hline\hline 
    \multicolumn{9}{|c|}{\sl 30 clusters} \\ \hline
    \system{MDAC($\vlambda^*, \vlambda_d$)}&36.3$^*$&56.9$^*$&51.3$^*$&33.2$^*$&43.4&89.4&51.8$^*$&-\\
     \system{DDS($\vlambda^*, \vlambda_d$)}&35.3&55.2&49.4&30.6&43.3&89.9&50.6&-\\ \hline
  \end{tabular}
  \caption{Non empty columns correspond to the target domains. For a given line, each non empty column correspond to one expert system. ($^*$)MDAC is significantly better than DDS.}
  \label{tab:unsupervised-da}
\end{table*}

\subsection{Unseen domains}\label{ssec:uda}
The top of Table \ref{tab:unsupervised-da} displays the performance on the unseen domain \domain{news} for NMT systems trained with the mixtures $\lambda^l \in \big[ \lambda_0, \lambda_{0.25}, \lambda_{0.5}, \lambda_{0.75}, \lambda_{1.0}\big]$ and dynamic data selection MDAC and DDS. These systems have insignificant differences in BLEU, showing that dynamic mixture does not improve the robustness of the NMT system against unseen domains. However, the fact that the performance of MDAC and DDS is close to the best performance guarantees the harmlessness of these algorithms to the robustness of the NMT model.

\subsection{Automatic clustering}\label{ssec:clda}
The bottom of Table \ref{tab:unsupervised-da} shows the performance of NMT systems adapted to each domain in using 30 clusters instead of 6 domains. According to Table \ref{tab:unsupervised-da}, MDAC underperforms the same version under the domain adaptation setting \ref{ssec:da}. However, automatic clustering could make the situations harder than the conventional settings where  
\section{Related Work \label{sec:related}}

Domain adaptation is an ancient and vexing problem that has been studied from many angles both for SMT and NMT. A survey of supervised and unsupervised DA for NMT is in \citet{Chu18asurvey}, where they distinguish between data-centric and model-centric DA, a view also adapted in the more recent review of \citet{Saunders21domain}. Our approach to DA in this paper clearly falls under the former category. We refer interested readers to these refrences, and do not discuss review DA any further.

Multi-domain adaptation (MDA) aims to develop systems that simultaneously bode well for several domains. Like for DA, techniques for supervised MDA typically combine one or several ingredients: (a) the specialization of data representations \citep{Kobus17domaincontrol} or of sub-networks \citep{Pham19generic} to differentiate the processing of each domain; (b) the use of adversarial techniques to neutralize differences between domains \cite{Britz17mixing,Zeng18multidomain}; (c) the use of automatic domain identification e.g.\ \citet{Jiang19multidomain}. Unsupervised MDA is considered in \citet{Farajian17multidomain} as an unsupervised DA problem.

Most approaches to adaptive / dynamic data selection take inspiration from \citet{Bengio09curriculum}, where the notion of \emph{curriculum learning} (CL) is initially introduced. CL is relies on a notion of ``easyness'' of a sample to schedule the presentation of the training data so that the easiest examples  presented first and the hardest, last. CL has been showed to speed up learning. While the initial aim was to improve and speed up training, it has also proven useful for domain adaptive / multi-domain / multi-lingual MT, based on alternative definitons of ``easyness''.

For instance, \citet{Zhang19curriculum} study supervised domain adaptation, and propose a curriculum approach which progressively extends the data used in training: in the early stages only in-data is used, while shards containing less relevant\footnote{Domain distance is computed with Lewis-Moore scores (based on the cross-entropy of in-domain LM).} data are introduced in later stages. This is somehow opposite to the recommendations of \citet{Vanderwees17dynamic}, whose \emph{gradual fine-tuning} progressively focuses on the in-domain data.\fyTodo{These have not been compared ? and also to what we do ?}

\citet{Kumar19reinforcement} use reinforcement techniques (deep Q-learning) to learn the curriculum strategy: in this study, complexity corresponds to difficulty levels and which are binned based on contrastive data selection. The reward is based on the increase of the development set loss that results from the actual data selection strategy.\fyTodo{Alert: what do we do during warm up ?} The same technique is recently applied to the training of a multilingual NMT system \citep{Kumar21learning}. \citet{Zhou20uncertainty} propose another curriculum-based approach which instead relies on \emph{instance uncertainty} as a measure of their difficulty and presents the data sample starting with the easiest (more predictable) first. Another contribution of this paper is an alternative criterium for stopping the training. More related to our problems, \citet{Wang20learning-multi} adapt curriculum learning for multi-domain adaptation, where an optimal instance weighting scheme is found using Bayesian optimization (derivative-free) techniques. Each step consists in (a) weighting the instances based on relevance features, (b) fine tuning current model using the (weighted) training set, and is applied iteratively to generate a sequence of models. The one that maximizes the development set performance is finally retained.

\section*{Acknowledgments}
\bibliographystyle{acl_natbib}
\bibliography{multidomain}
%\appendix
\end{document}